Mapping out the punishment strategy space

### 2023-10-20 report.Rmd
starting writing manuscript

We can also look at the most common behavioural patterns across all six games (twelve punishment decisions in total). Below, we summarise the five most common patterns of behaviour as strings of twelve 0s and 1s, indicating whether participants did (1) or did not (0) punish in each game, with a brief explanation of this pattern and its frequency in the dataset.

```{r patterns}
kable(tar_read(tablePatterns), booktabs = TRUE)
```

@@

We can also estimate the frequencies of different strategies using a Bayesian approach. We construct a model that contains our _apriori_ predictions for the different punishment strategies, and the feed the model our raw data to estimate the relative probabilities of following each strategy. In the model, we assume that participants sometimes make errors in converting their strategy into behaviour (5% error rate), which could explain why many of the participants were unable to be classified into a strategy type in our raw counts above. The Bayesian model is fitted in the probabilistic programming language Stan.

```{r post probs}
P <- tar_read(post1)$alpha
for (i in 1:nrow(P)) P[i,] <- rethinking::softmax(P[i,])
anti <- P[,10]
```

In line with our raw counts, the model predicts that participants follow the "anti-punish" strategy with a probability of `r round(median(anti), 2)`, 95% credible interval [`r round(quantile(anti, 0.025), 2)` `r round(quantile(anti, 0.975), 2)`]. We plot the posterior probabilities for the remaining punishment strategies below.

```{r model1, message=FALSE, out.width="80%", fig.align='center'}
tar_read(plot1)
```

Taking advantage of all available data, the model suggests that "egalitarian" is the most common punishment strategy. All other strategies have median posterior probabilities less than 10%. In order of decreasing probability, the next most common punishment strategies are "norm-enforcing", "seek advantageous inequity", "deterrent", and "avoid disadvantageous inequity", although all are similarly likely. "Competitive" and "antisocial" punishment strategies are the most unlikely.

@@

We then fit a series of models predicting the usage of different strategies from a variety of predictors. In each plot that follows, we show the posterior predictions from a model that includes a single variable as a predictor of all ten strategies simultaneously. The plots show the median regression lines with shaded 95% credible intervals. The model-estimated slopes for each strategy are included in each panel.

```{r plot2, fig.width=7, fig.height=4, fig.align='center', out.width="100%"}
tar_read(plot2_Age)
tar_read(plot2_SelfRate_1)
tar_read(plot2_SelfRate_2)
tar_read(plot2_SelfRate_3)
tar_read(plot2_SelfRate_4)
tar_read(plot2_SelfRate_5)
tar_read(plot2_SelfRate_6)
tar_read(plot2_SelfRate_7)
tar_read(plot2_SelfRate_8)
tar_read(plot2_SelfRate_9)
tar_read(plot2_SelfRate_10)
tar_read(plot2_SelfRate_11)
```

>>>>

We assume here that people are following a particular strategy exactly and are not comitting any errors or mistakes. However, it is possible that some participants intended to follow a particular strategy, but made a mistake on a particular decision. We repeat the counts and proportions in the table above, but include behavioural patterns that have one mistake from the exact strategy (note that the proportions will no longer sum to 1).

```{r strategyCountsMistake1}
kable(tar_read(tableStrategyCounts_Mistake1), booktabs = TRUE)
```

And including behavioural patterns that have one or two mistakes.

```{r strategyCountsMistake2}
kable(tar_read(tableStrategyCounts_Mistake2), booktabs = TRUE)
```

And finally, including behavioural patterns that have one, two, or three mistakes.

```{r strategyCountsMistake3}
kable(tar_read(tableStrategyCounts_Mistake3), booktabs = TRUE)
```

We can also look at the most common behavioural patterns across all six games (twelve punishment decisions in total). Below, we summarise the 25 most common patterns of behaviour as strings of twelve 0s and 1s, indicating whether participants did (1) or did not (0) punish in each game, with a brief explanation of this pattern and its frequency in the dataset.

```{r patternsUK}
kable(tar_read(tablePatterns), booktabs = TRUE)
```

These behavioural patterns can be visualised, along with their frequency of usage, on a splits graph. This graph plots the distance between the difference strategies as proportional to the number of substitutions required to get from one to another. The top of the graph captures the less punitive strategies, and the strategies become more punitive towards the bottom of the graph. We only include behavioural patterns followed by at least three participants.

```{r splitsGraph, fig.width=7, fig.height=7}
tar_read(plotSplits)
```

@@

We can also estimate the frequencies of different strategies using a Bayesian approach. We construct a model that contains our _apriori_ predictions for the different punishment strategies, and the feed the model our raw data to estimate the relative probabilities of following each strategy. In the model, we assume that participants sometimes make errors in converting their strategy into behaviour (5% error rate), which could explain why many of the participants were unable to be classified into a strategy type in our raw counts and proportions above. The Bayesian model is fitted in the probabilistic programming language Stan.

```{r model1, message=FALSE, error=FALSE, warning=FALSE, out.width="100%"}
tar_read(plot1.1)
```

The pattern is similar in both countries. Taking advantage of all available data, the model suggests that "anti-punish" is the most common strategy. Of the punishment strategies, "egalitarian" is the most common. All other strategies have median posterior probabilities less than 10%. "Competitive" and "antisocial" punishment strategies are the most unlikely.

Why is egalitarian the winner in this model, considering that it wasn't an obvious winner in the raw counts? One explanation is that many participants punished only in the third-party punishment game (see table above). The model does not include this pattern of behaviour as an explicit strategy. This pattern of behaviour is one substitution away from egalitarian, but two substitutions away from avoid DI, so it upweights egalitarian accordingly.

@@

We then fit a series of models predicting the usage of different strategies from a variety of predictors. In each plot that follows, we show the posterior slopes from several models that include different variables as predictors of all ten strategies in both countries. Each predictor is included separately, apart from SDO and RWA where we include both in the model simultaneously to control for one another.

## Demographics

```{r echo=F}
tar_read(plotDemsNoTPP)
```

\newpage

## Self-ratings

```{r echo=F}
tar_read(plotSlidersNoTPP)
```

\newpage

## Big-6 personality and SVO

```{r echo=F}
tar_read(plotPersNoTPP)
```

\newpage

## Politics and religion

```{r echo=F}
tar_read(plotPolRelNoTPP)
```
